---
title: "Data Cleaning Code"
author: "Lucy Moctezuma"
date: "2023-04-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/lucyl/Desktop/LEARNING/#00_MASTERS/STAT 632 - Linear and Logistic Regression/Final Project")
library(tidyverse)
library(dplyr)
```

# Loading Dataset and Data cleaning

## Canthril Ladder Score
```{r}
hp_df <-read.csv("happiness-cantril-ladder.csv", header=T)
hp_df<-hp_df[which(hp_df$Year==2019),]
hp_df <- subset(hp_df, select = -c(Code,Year))
colnames(hp_df)[2]="lfsa"
head(hp_df)
```
## Life Expectancy
```{r}
le_df <-read.csv("life-expectancy.csv", header=T)
le_df<-le_df[which(le_df$Year==2019),]
le_df <- subset(le_df, select = -c(Code,Year))
colnames(le_df)[2]="lfex"
head(le_df)
```
## Internet usage
```{r}
int_df <-read.csv("share-of-individuals-using-the-internet.csv", header=T)
int_df<-int_df[which(int_df$Year==2019),]
int_df <- subset(int_df, select = -c(Code,Year))
colnames(int_df)[2]="inet"
head(int_df)
```
## Mental Health Prevalence
```{r}
mh_df <-read.csv("prevalence-by-mental-and-substance-use-disorder.csv", header=T)
mh_df <-mh_df[which(mh_df$Year==2019),]
colnames(mh_df)[4]="schz"
colnames(mh_df)[5]="bipd"
colnames(mh_df)[6]="eatd"
colnames(mh_df)[7]="anxi"
colnames(mh_df)[8]="drgu"
colnames(mh_df)[9]="depr"
colnames(mh_df)[10]="aud"
mh_df <- subset(mh_df, select = -c(Code,Year))
head(mh_df)
```
## Economic Inequality
```{r}
ineq_df <-read.csv("economic-inequality-gini-index.csv", header=T)
# renaming Argentina - urban to just Argentina
ineq_df[ineq_df$Entity=="Argentina - urban","Entity"] <- "Argentina"

# deleting repeated countries where the values are subdivided into urban and rural areas (only China, India and Indonesia)
ineq_df<- ineq_df[!grepl("rural", ineq_df$Entity),]
ineq_df<- ineq_df[!grepl("urban", ineq_df$Entity),]

# Choosing all gini index measures before 2020 
ineq_df<- ineq_df[which(ineq_df$Year < 2020),]
head(ineq_df)
```
```{r}
# Choosing the last measure of gini index done since 2019
ineq_df1 <- ineq_df[order(ineq_df$Year, decreasing = TRUE), ] 
ineq_df1 <- Reduce(rbind,                                 
                   # Top N highest values by country
                    by(ineq_df1,
                       ineq_df1["Entity"],
                       head,
                       n = 1))
# Erasing country codes 
ineq_df1 <- subset(ineq_df1, select = -c(Code))

# Changing to easier column names
colnames(ineq_df1)[2]="year"
colnames(ineq_df1)[3]="gini"

head(ineq_df1)
```

## Suicide Rates by Age group
```{r}
suic_df <- read.csv("suicide-rates-by-age-detailed.csv", header=T)
suic_df <- suic_df[which(suic_df$Year == 2019),]
suic_df <- subset(suic_df, select = -c(Code,Year))
colnames(suic_df)[2] = "suic.70plus"
colnames(suic_df)[3] = "suic.50to69"
colnames(suic_df)[4] = "suic.allages"
colnames(suic_df)[5] = "suic.5to14"
colnames(suic_df)[6] = "suic.15to49"

suic_df <- suic_df[,c(1,4,5,6,3,2)]
head(suic_df)
```
## Homicide Rates
```{r}
hom_df <- read.csv("homicide-rate.csv", header=T)
hom_df <- hom_df[which(hom_df$Year == 2019),]
hom_df <- subset(hom_df, select = -c(Code,Year))
colnames(hom_df)[2] = "homc"
head(hom_df)
```
# Merging all dataframes
```{r}
# merging all data frames except for income inequality by Entity
df_list <- list(hp_df,le_df,int_df, mh_df,suic_df, hom_df) 
alldata <- df_list %>% reduce(inner_join, by='Entity')

# merging data frames with income inequality
df <- merge(x=alldata,y=ineq_df1, by="Entity", all.x=T)
head(df)

```
## Adding Regional Variable

```{r}
afr <- c("Algeria", "Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cameroon", "Cape Verde", "Central African Republic", "Chad", "Comoros", "Cote d'Ivoire", "Democratic Republic of the Congo", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Republic of the Congo", "Rwanda", "São Tomé and Príncipe", "Senegal", "Seychelles", "Sierra Leone", "South Africa", "South Sudan", "Eswatini", "Togo", "Uganda", "Tanzania", "Zambia", "Zimbabwe")

amr <- c("Antigua and Barbuda", "Argentina", "Bahamas", "Barbados", "Belize", "Bolivia", "Brazil", "Canada", "Chile", "Colombia", "Costa Rica", "Cuba", "Dominica", "Dominican Republic", "Ecuador", "El Salvador", "Grenada", "Guatemala", "Guyana", "Haiti", "Honduras", "Jamaica", "Mexico", "Nicaragua", "Panama", "Paraguay", "Peru", "Saint Kitts and Nevis", "Saint Lucia", "Saint Vincent and the Grenadines", "Suriname", "Trinidad and Tobago", "United States", "Uruguay", "Venezuela")

sear <- c("Bangladesh", "Bhutan", "North Korea", "India", "Indonesia", "Maldives", "Myanmar", "Nepal", "Sri Lanka", "Thailand", "Timor-Leste")

eur <- c("Albania", "Andorra", "Armenia", "Austria", "Azerbaijan", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria", "Croatia", "Cyprus", "Czech Republic","Czechia", "Denmark", "Estonia", "Finland", "France", "Georgia", "Germany", "Greece", "Hungary", "Iceland", "Ireland", "Israel", "Italy", "Kazakhstan", "Kyrgyzstan", "Latvia", "Lithuania", "Luxembourg", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia", "Norway", "Poland", "Portugal", "Romania", "Russia", "San Marino", "Serbia", "Slovakia", "Slovenia", "Spain", "Sweden", "Switzerland", "Tajikistan", "Turkey", "Turkmenistan", "Ukraine", "United Kingdom", "Uzbekistan")

emr <- c("Afghanistan", "Bahrain", "Djibouti", "Egypt", "Iran", "Iraq", "Jordan", "Kuwait", "Lebanon", "Libya", "Morocco", "Oman", "Pakistan", "Qatar", "Saudi Arabia", "Somalia", "Sudan", "Syria", "Tunisia", "United Arab Emirates", "Yemen", "Palestine")

wpr <- c("Australia", "Brunei", "Cambodia", "China", "Cook Islands", "Fiji", "Japan", "Kiribati", "Laos", "Malaysia", "Marshall Islands", "Micronesia", "Mongolia", "Nauru", "New Zealand", "Niue", "Palau", "Papua New Guinea", "Philippines", "Samoa", "Singapore", "Solomon Islands", "South Korea", "Tonga", "Tuvalu", "Vanuatu", "Vietnam")

df$region <- c()
for (i in 1:nrow(df)){
  if (df$Entity[i] %in% afr){
    df$region[i] <- "AFR"
  }else if (df$Entity[i] %in% amr){
    df$region[i] <- "AMR"
  }else if (df$Entity[i] %in% sear){
    df$region[i] <- "SEAR"
  }else if (df$Entity[i] %in% eur){
    df$region[i] <- "EUR"
  }else if (df$Entity[i] %in% emr){
    df$region[i] <- "EMR"
  }else if (df$Entity[i] %in% wpr){
    df$region[i] <- "WPR"
  }else{
    df$region[i] <- "NOT CODED"
  }
}
df$region <- as.factor(df$region)
```
# Checking for missing values
```{r}
df[is.na(df$gini),]
```
There are only 6 countries with missing gini index values. I have decided to use their regional mean gini to fill up their values. I will now proceed to create a regional variable. For example I will check other countries from (emr) region that we do have their gini indexes, take their mean and then input that on (Kuwait, Saudi Arabia , Bahrain and Afghanistan), since they belong to that region.  

## Dealing with missing data gini values and Year
```{r}
# getting emr_gini to replace missing gini values in the countries (Afghanistan, Bahrain, Kuwait and Saudi Arabia)
emr_f<- df[which(df$region == "EMR"),] %>% drop_na()
emr_gini<-mean(emr_f$gini)
df[df$Entity=="Afghanistan","gini"] <- emr_gini
df[df$Entity=="Bahrain","gini"] <- emr_gini
df[df$Entity=="Kuwait","gini"] <- emr_gini
df[df$Entity=="Saudi Arabia","gini"] <- emr_gini

# getting wpr_gini to replace missing gini values in the countries (Singapore and New Zealand) 
wpr_f<- df[which(df$region == "WPR"),] %>% drop_na()
wpr_gini<-mean(wpr_f$gini)
df[df$Entity=="New Zealand","gini"] <- wpr_gini
df[df$Entity=="Singapore","gini"] <- wpr_gini
```
## Looking at all the data once more
```{r}
head(df)
```
## Writting csv file
```{r}
write.csv(df, "C:/Users/lucyl/Desktop/LEARNING/#00_MASTERS/STAT 632 - Linear and Logistic Regression/Final Project/merged_df.csv",row.names = F)
```
