---
title: "Lab Name"
author: "Author Name"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r load-packages, message=FALSE}
# Set up directory
setwd("C:/Users/lucyl/Desktop/LEARNING/#00_MASTERS/STAT 632 - Linear and Logistic Regression/Final Project/Final Project Paper")
library(tidyverse)
library(openintro)
library(emmeans)
library(ggplot2)
library(car)
library(dplyr)
library(devtools)
library(stringr)
library(caret)

# Making pretty plots
library(RColorBrewer)
library(kableExtra)
library(rworldmap)
library(ggthemes)
library(sf)
library(ggfortify)
library(grid)

# libraries for regular plots
library(ggplot2)
library(GGally)
library(tidyverse)
library(gridExtra)
library(ggpubr)
library(ggcorrplot)

```

# Loading Data

```{r code-chunk-label}
# load and view data
df <- read.csv(file ="merged_df.csv",header=TRUE)
#Make country names into index now and subsetting
data <- subset(df, select=-c(Entity, year))
rownames(data) <- df$Entity
data$region <- as.factor(data$region)
head(data)
```
```{r}
summary(data)
```
# REASONS why we chose to ONLY Interpret the model before stepwise regression

- Smith, G. Step away from Stepwise. J Big Data 5, 32 (2018). https://doi.org/10.1186/s40537-018-0143-6
- Slide Lectures from Dr. Kerr as well.
- We used the stepwise regression only to select the best predictive model.

# Additive model (Some transformations)
```{r}
addmodel<- lm(lfsa ~ lfex + inet + log(schz)+ bipd + log(eatd) + anxi + log(drgu) + depr + aud + log(suic.5to14) + log(suic.15to49) + suic.50to69 + log(suic.70plus) + log(homc) + gini + region, data = data)
summary(addmodel)

addmodel_std<- lm(lfsa ~ scale(lfex) + scale(inet) + scale(log(schz))+ scale(bipd) + scale(log(eatd)) + scale(anxi) + scale(log(drgu)) + scale(depr) + scale(aud) + scale(log(suic.5to14)) + scale(log(suic.15to49)) + scale(suic.50to69) + scale(log(suic.70plus)) + scale(log(homc)) + scale(gini) + region, data = data)
summary(addmodel_std)
```
# Check model if you only included the the significant predictors
```{r}
addmodel_try<- lm(formula = lfsa ~ (inet + log(eatd) + log(suic.5to14) + suic.50to69 + gini + region), data = data)
summary(addmodel_try)
```

## Assumption Checking

```{r}
# Checking residual plots
par(mfrow=c(1,2))
plot(addmodel, which = c(1,2)) 
shapiro.test(rstandard(addmodel))
```
The graphs below are the ones that prompted the transformations, there
are the residuals versus the variables with transformation.

```{r}
# checking residuals against each predictor
# standard residual vs each measure (lfex, inet, homc , gini)
par(mfrow=c(2,2),mar=c(4.5, 4.5, 2, 2))
plot(rstandard(addmodel)~lfex, data=data)
plot(rstandard(addmodel)~inet, data=data)
plot(rstandard(addmodel)~log(homc), data=data)
plot(rstandard(addmodel)~gini, data=data)

# standard residual vs each mental health measure
par(mfrow=c(2,2),mar=c(4.5, 4.5, 2, 2))
plot(rstandard(addmodel)~log(schz), data=data)
plot(rstandard(addmodel)~bipd, data=data)
plot(rstandard(addmodel)~log(eatd), data=data)
plot(rstandard(addmodel)~anxi, data=data)

par(mfrow=c(2,2),mar=c(4.5, 4.5, 2, 2))
plot(rstandard(addmodel)~log(drgu), data=data)
plot(rstandard(addmodel)~depr, data=data)
plot(rstandard(addmodel)~aud, data=data)

# standard residual vs each suicide measure
par(mfrow=c(2,2),mar=c(4.5, 4.5, 2, 2))
plot(rstandard(addmodel)~log(suic.5to14), data=data)
plot(rstandard(addmodel)~log(suic.15to49), data=data)
plot(rstandard(addmodel)~suic.50to69, data=data)
plot(rstandard(addmodel)~log(suic.70plus), data=data)

# standard residual vs region (not for assumption checking)
par(mfrow=c(1,1),mar=c(4.5, 4.5, 2, 2))
plot(rstandard(addmodel)~region, data=data)
```

```{r}
# Checking leverage and outliers
plot(addmodel, which =5)
abline(v=(2*21)/nrow(data),col="blue", lty=2)
abline(h=c(-2,2),col="red", lty=2)
```

```{r}
# Identifying outliers and high leverage points
data[which(hatvalues(addmodel)>0.4),]
data[which(abs(rstandard(addmodel))>2.2),]
```

```{r}
# Checking multicollinearity issues
vif(addmodel)
```

```{r}
boxCox(addmodel)
summary(powerTransform(addmodel))
```
## Interpretations for rescaled Additive Model:
Below are hypotheses:

$H_0: All \ \beta_i = 0, where \ i= 1...19$

$H_a: At \ least \ one \ \beta_i \neq 0$

At $\alpha = 0.05$ the only significant predictors we found were:

inet              0.015583   0.005813   2.681 0.008475 **
    
log(eatd)         0.784008   0.266330   2.944 0.003957 **

regionEMR        -0.844344   0.322149  -2.621 0.010008 *

We did found some statistically significant predictor variables in our additive model, mainly internet and log(eatdisorder) were positively associated with Cantril ladder score.
We saw that regionEMR was negatively associated with the Cantril ladder score, and had the highest absolute value in terms of coefficient.

## Anova (Null versus Additive model)

```{r}
null_model<- lm(lfsa~1, data=data)
anova(null_model, addmodel)
```
This is essentially the F test showing that yes at least one of the
coefficients for each of our variables is significantly not zero.

# GINI interaction model

```{r}
gini_model<- lm(lfsa ~ (lfex + inet + log(schz)+ bipd + log(eatd) + anxi + log(drgu) + depr + aud + log(suic.5to14) + log(suic.15to49) + suic.50to69 + log(suic.70plus) + log(homc) + gini + region)*gini, data = data)  

gini_model_std <- lm(lfsa ~ (scale(lfex) + scale(inet) + scale(log(schz))+ scale(bipd) + scale(log(eatd)) + scale(anxi) + scale(log(drgu)) + scale(depr) + scale(aud) + scale(log(suic.5to14)) + scale(log(suic.15to49)) + scale(suic.50to69) + scale(log(suic.70plus)) + scale(log(homc)) + scale(gini) + region)*scale(gini), data = data)  

summary(gini_model)

```
## Assumption Checking
```{r}
# Checking residual plots
par(mfrow=c(1,2))
plot(gini_model_std, which = c(1,2)) 
shapiro.test(rstandard(gini_model_std))
```
### GINI Interaction model (Post Stepwise)
```{r}
gini_model2<-step(gini_model, trace=0)
summary(gini_model2)
```
### Assumptions GINI_Interaction Model (Post Stepwise):
```{r}
par(mfrow=c(1,2))
plot(gini_model2, which = c(1,2)) 
shapiro.test(rstandard(gini_model2))
```
# Conclusions for GINI_Interaction Model (Post Stepwise)
We found the following interactions to be significant:
log(eatd):gini        -9.01008    2.23247  -4.036 9.83e-05 ***
log(suic.15to49):gini -3.06569    1.46168  -2.097 0.038151 *  
log(suic.70plus):gini -4.00327    1.34571  -2.975 0.003573 ** 

I was particularly curious as to why log(eatd) was significant in our additive model was significant and wanted to see if this particular variable interacted with some sort of economic measure such as gini, interestingly it did, and it actually had the highest absolute value of coefficient. (Their effects had an inverse relationship) 

# INET interaction model
```{r}
inet_model<- lm(lfsa ~ (lfex + inet + log(schz)+ bipd + log(eatd) + anxi + log(drgu) + depr + aud + log(suic.5to14) + log(suic.15to49) + suic.50to69 + log(suic.70plus) + log(homc) + gini + region)*inet, data = data)
summary(inet_model) 

inet_model_std<- lm(lfsa ~ (scale(lfex) + scale(inet) + scale(log(schz))+ scale(bipd) + scale(log(eatd)) + scale(anxi) + scale(log(drgu)) + scale(depr) + scale(aud) + scale(log(suic.5to14)) +scale(log(suic.15to49)) + scale(suic.50to69) + scale(suic.70plus) + scale(log(homc)) + scale(gini) + region)*scale(inet), data = data) 
summary(inet_model_std) 
```
## Assumption Checking
```{r}
par(mfrow=c(1,2))
plot(inet_model, which = c(1,2)) 
shapiro.test(rstandard(inet_model))
```
### INET Interaction model (Post Stepwise)

```{r}
inet_model2<-step(inet_model, trace=0)
summary(inet_model2)
```
### Assumptions INET_Interaction Model (Post Stepwise):

```{r}
par(mfrow=c(1,2))
plot(inet_model2, which = c(1,2)) 
shapiro.test(rstandard(inet_model2))
```
# Conclusions for INET_Interaction Model (Post Stepwise)
We found the following interactions to be significant:
inet:bipd              0.071839   0.021844   3.289  0.00137 ** 
inet:log(suic.15to49)  0.009527   0.004349   2.191  0.03068 *  
inet:regionSEAR        0.042109   0.015355   2.742  0.00717 ** 

Interestingly we found that they had a positive interaction effect with internet, meaning the higher the effect on one variable, the higher the effect the other variable had on the Cantril ladder score.

# DRUG interaction model
```{r}
drgu_model<- lm(lfsa ~ (lfex + inet + log(schz)+ bipd + log(eatd) + anxi + log(drgu) + depr + aud + log(suic.5to14) +log(suic.15to49) + suic.50to69 + log(suic.70plus) + log(homc) + gini + region)*log(drgu), data = data)  
summary(drgu_model)

drgu_model_std<- lm(lfsa ~ (scale(lfex) + scale(inet) + scale(log(schz))+ scale(bipd) + scale(log(eatd)) + scale(anxi) + scale(log(drgu)) + scale(depr) + scale(aud) + scale(log(suic.5to14)) + scale(log(suic.15to49)) + scale(suic.50to69) + scale(suic.70plus) + scale(log(homc)) + scale(gini) + region)*scale(log(drgu)), data = data) 

summary(drgu_model_std)
```
## Assumption Checking
```{r}
par(mfrow=c(1,2))
plot(drgu_model, which = c(1,2)) 
shapiro.test(rstandard(drgu_model))
```
### DRUG Interaction model (Post Stepwise)

```{r}
drgu_model2<-step(drgu_model, trace=0)
summary(drgu_model2)
```
### Assumptions DRUG_Interaction Model (Post Stepwise):
```{r}
par(mfrow=c(1,2))
plot(drgu_model2, which = c(1,2)) 
shapiro.test(rstandard(drgu_model2))
```
# Conclusions for DRGU_Interaction Model (Post Stepwise)
We found the following interactions to be significant:
anxi:log(drgu)         0.442245   0.119216   3.710 0.000331 ***
log(drgu):depr        -0.326252   0.148510  -2.197 0.030189 *  
log(drgu):suic.70plus -0.020769   0.009910  -2.096 0.038456 *  
log(drgu):regionAMR   -1.746800   0.695426  -2.512 0.013504 *  
log(drgu):regionEMR   -1.681711   0.846370  -1.987 0.049482 *  Barely significant
log(drgu):regionWPR   -2.185510   0.793040  -2.756 0.006883 ** 

We found the most number of interactions to be significant in the log(drgu) interaction model one, with mostly negative interactions coefficients. For the interactions with regions we found that when log(drug) increases, the effect of the regions (AMR, EMR, WPR) decreases on average in respects to the mean score of Cantril ladder for the AFR region. 

# Comparing all 3 (Post Stepwise) interaction models

```{r}
AIC(addmodel,gini_model2, inet_model2,drgu_model2)
```

# Plotting residuals using ggplot

```{r}
#Assumption checking for plots
r1<-autoplot(addmodel, which = 1:2, ncol = 2, label.repel = TRUE, colour = "grey26")+theme_bw()+theme(text = element_text(size=10), legend.position = "none")+ylab("Standarized residuals")
r1<-grid.arrange(r1, grobs = r1@plots, top = textGrob(expression(bold(underline("Additive Model Assumptions"))), gp=gpar(col="grey26")),widths=c(1,1))
r1
ggsave("addmodel_residplots.jpeg",r1, bg="white")
```

```{r}
#Assumption checking for plots
r1.1<-autoplot(gini_model2, which = 1:2, ncol = 2, label.repel = TRUE, colour ="seagreen")+theme_bw()+theme(text = element_text(size=10), legend.position = "none")+ylab("Standarized residuals")
r1.1<-grid.arrange(r1.1, grobs = r1.1@plots, top = textGrob(expression(bold(underline("Gini Interaction Model Assumptions"))), gp=gpar(col="seagreen")),widths=c(1,1))

r1.1
ggsave("gini_inter_residplots.jpeg",r1.1, bg="white")
```

```{r}
#Assumption checking for plots
r1.2<-autoplot(inet_model2, which = 1:2, ncol = 2, label.repel = TRUE, colour="steelblue")+theme_bw()+theme(text = element_text(size=10), legend.position = "none")+ylab("Standarized residuals")
r1.2<-grid.arrange(r1.2, grobs = r1.2@plots, top = textGrob(expression(bold(underline("Internet Usage Interaction Model Assumptions"))), gp=gpar(col="steelblue")),widths=c(1,1))
r1.2
ggsave("inet_inter_residplots.jpeg",r1.2, bg="white")
```

```{r}
#Assumption checking for plots
r1.3<-autoplot(drgu_model2, which = 1:2, ncol = 2, label.repel = TRUE, colour="maroon4")+theme_bw()+theme(text = element_text(size=10), legend.position = "none")+ylab("Standarized residuals")
r1.3<-grid.arrange(r1.3, grobs = r1.3@plots, top = textGrob(expression(bold(underline("Drug Usage Interaction Model Assumptions"))), gp=gpar(col="maroon4")),widths=c(1,1))
r1.3
ggsave("drgu_inter_residplots.jpeg",r1.3, bg="white")
```
## Making predictions using our best AIC model
```{r}
pred <- addmodel %>% predict(data[,-1])
add_df<-data.frame( R2 = R2(pred, data$lfsa),
RMSE = RMSE(pred, data$lfsa),
MAE = MAE(pred, data$lfsa))

pred_inet <- inet_model2 %>% predict(data[,-1])
i_df<-data.frame( R2 = R2(pred_inet, data$lfsa),
RMSE = RMSE(pred_inet, data$lfsa),
MAE = MAE(pred_inet, data$lfsa))

pred_gini <- gini_model2 %>% predict(data[,-1])
g_df<- data.frame( R2 = R2(pred_gini, data$lfsa),
RMSE = RMSE(pred_gini, data$lfsa),
MAE = MAE(pred_gini, data$lfsa))

pred_drgu <- drgu_model2 %>% predict(data[,-1])
d_df<- data.frame( R2 = R2(pred_drgu, data$lfsa),
RMSE = RMSE(pred_drgu, data$lfsa),
MAE = MAE(pred_drgu, data$lfsa))

total <- rbind(add_df,i_df, g_df, d_df)
rownames(total) <- c("Additive model","Internet_model","Gini_model","DrugUse_model")
total
```

